\section{A Further Optimized Algorithm for DHL$(\circ)$ Materialization}
\label{sec:practicalAlg}

In this section, we first discuss that Algorithm~$\mathsf{A}_{opt}$ can hardly work in practice.
In order to make Algorithm~$\mathsf{A}_{opt}$ more practical,
we propose to modify Algorithm~$\mathsf{A}_{opt}$ and give an algorithm variant Algorithm~$\mathsf{A}_{prc}$.
We show that Algorithm~$\mathsf{A}_{prc}$ can also be restricted to an NC version
when materializing DHL$(\circ)$ ontologies that follow both of the simple-concept
and the role-concept restrictions.

\subsection{Reducing Computing Space}

In previous sections, Algorithm~$\mathsf{A}_{opt}$ is mainly used for theoretical analysis.
However, this algorithm can hardly work in practice due to its inherently
high requirement of computing space. Specifically,
Algorithm~$\mathsf{A}_{opt}$ constructs a materialization graph by checking all possible rule
instantiations in $P^*$ where $P$ is a datalog program that corresponds to an ontology.
One can check that $|P^*|$ could be the square or cube of the number of constants occurring in $P$.
Consider a datalog rule of the form `$R(x,y),A(y)\rightarrow B(x)$'
that is rewritten from the axiom of the form $\exists R.A\sqsubseteq B$;
since there are two variables in this rule, the number of all possible rule instantiations
is $|\textbf{IN}|^2$; if there are
1000 individuals in \textbf{IN}, this number is 1 million.
Similarly, for a datalog rule rewritten from an axiom of the form (R3) or (R4),
the number of all possible rule instantiations is $|\textbf{IN}|^3$.
It is no doubt that a plain implementation of Algorithm~$\mathsf{A}_{opt}$ would be delayed
when the target datalog program or ontology
tends to be large in size.
On the other hand, from Examples~\ref{exp:mg}, \ref{exp:dllite} and \ref{exp:dhl},
we can observe that the rule instantiations used for the construction
of materialization graphs always cover a small part of $P^*$ with respect to the target datalog program $P$.
Thus, we consider reducing the computing space of Algorithm~$\mathsf{A}_{opt}$
by narrowing down the scope of rule instantiations to be checked.
Our strategy is to restrict that, \emph{in each iteration of Algorithm~$\mathsf{A}_{opt}$,
each of the checked rule instantiations should involve at least one body atom that has been
added to the constructed materialization graph}.
Since DHL$(\circ)$
is our focus, we explain how to apply the above strategy in DHL$(\circ)$ materialization
as follows.

We first consider a datalog rule of the
form `$A(x)\rightarrow B(x)$' that corresponds to (T1). The above strategy requires that,
in each iteration, Algorithm~$\mathsf{A}_{opt}$ can only check the rule instantiations
of the form `$A(a)\rightarrow B(a)$' where $A(a)$ has been added to the constructed materialization
graph $\mathcal{G}$. If there are $n$ atoms of the form $A(x)$ that have been added to $\mathcal{G}$,
the number of all checked rule instantiations of the above rule is also $n$, instead of $|\textbf{IN}|$.
This is because that, for each assertion $A(a)$,
the datalog rule `$A(x)\rightarrow B(x)$'
has only one rule instantiation `$A(a)\rightarrow B(a)$' where
the variable $x$ is substituted by the constant $a$.
The cases of (R1) and (R2) can be analyzed similarly.
We now consider the datalog rules that have two body atoms, i.e.,
the datalog rules of the forms `$R(x,y),A(y)\rightarrow B(x)$' (see (T3)),
`$A_1(x),A_2(x)\rightarrow B(x)$' (see (T2)) and `$R_1(x,y),R_2(y,z)\rightarrow R(x,z)$' (see (R3) and (R4)).
We require that, for each rule instantiation of these rules checked by Algorithm~$\mathsf{A}_{opt}$,
at least one body atom has been added to the constructed materialization
graph $\mathcal{G}$; thus, it can be checked that, in each iteration of Algorithm~$\mathsf{A}_{opt}$,
the number of checked rule instantiations
is at most $k|\textbf{IN}|$ where $k$ is the number of atoms that have been
added to $\mathcal{G}$.
Note that, for rule instantiations of two body atoms,
we do not require that both of these two body atoms have been added to
the constructed materialization graph.
Otherwise, the algorithm would
perform as the same as Algorithm~$\mathsf{A}_{bsc}$, and thus, the
optimizations used in Algorithm~$\mathsf{A}_{opt}$ cannot further work.



\subsection{Further Optimizing Algorithm~$\mathsf{A}_{opt}$}


We use the above method of narrowing down the scope of rule instantiations to
modify Algorithm~$\mathsf{A}_{opt}$ and obtain an algorithm variant Algorithm~$\mathsf{A}_{prc}$.
Recall that,
in each iteration of \ref{alg3:updateG}, Algorithm~$\mathsf{A}_{opt}$
checks all possible rule instantiations and computes a \texttt{rch} relation
and its transitive closure to determine the existence of SWD paths.
We let Algorithm~$\mathsf{A}_{prc}$ conduct the same work with
narrowing down the scope of rule instantiations to be checked as well. This can be described
by the following algorithm.\\

\noindent\texttt{Algorithm~$\mathsf{PRC}$}. This algorithm has inputs (1)
a DHL$(\circ)$ ontology $\mathcal{O}=\langle\mathcal{T},\mathcal{R},\mathcal{A}\rangle$
and its datalog program $P=\langle R, \textbf{I}\rangle$;
(2) a (partial) materialization graph $\mathcal{G}=\langle V,E\rangle$ that is constructed from $P$.
This algorithm outputs a \texttt{rch} relation $S_{\textit{\tiny rch}}$ that
is computed as follows:

\begin{enumerate}[leftmargin=4ex,label=$\bullet$]
\item add \texttt{rch}$(A(a),B(a))$ to $S_{\textit{\tiny rch}}$ where $A(a)\rightarrow B(a)\in P^*$, $\mathcal{O}\models A\sqsubseteq_* B$ and $A(a)\in V$;

\item add \texttt{rch}$(A(b),B(a))$ to $S_{\textit{\tiny rch}}$ where $R(a,b),A(b)\rightarrow B(a)\in P^*$, $\exists R.A\sqsubseteq B\in\mathcal{T}$ and $R(a,b)\in V$;

\item add \texttt{rch}$(A_2(a),B(a))$ to $S_{\textit{\tiny rch}}$ where $A_1(a),A_2(a)\rightarrow B(a)\in P^*$,
    $A_1\sqcap A_2\sqsubseteq B\in\mathcal{T}$ and $A_1(a)\in V$;

\item add \texttt{rch}$(A_1(a),B(a))$ to $S_{\textit{\tiny rch}}$ where $A_1(a),A_2(a)\rightarrow B(a)\in P^*$,
    $A_1\sqcap A_2\sqsubseteq B\in\mathcal{T}$ and $A_2(a)\in V$;

\item add \texttt{rch}$(R(a,b)$, $S(a,b))$ to $S_{\textit{\tiny rch}}$ where $R(a,b)\rightarrow S(a,b)\in P^*$,
    $\mathcal{O}\models R\sqsubseteq_* S$ and $R(a,b)\in V$;

\item add \texttt{rch}$(R(a,b)$, $S(b,a))$ to $S_{\textit{\tiny rch}}$ where $R(a,b)\rightarrow S(b,a)\in P^*$,
    $\mathcal{O}\models R\sqsubseteq_* S^-$ and $R(a,b)\in V$;

\item add \texttt{rch}$(R_2(b,c),R_3(a,c))$ to $S_{\textit{\tiny rch}}$ where $R_1(a,b),R_2(b,c)\rightarrow R_3(a,c)\in P^*$,
    $R_1\circ R_2\sqsubseteq R_3\in\mathcal{R}$ (the case
    where $R_1\equiv R_2\equiv R_3$ is also involved) and $R_1(a,b)\in V$;

\item add \texttt{rch}$(R_1(a,b)$, $R_3(a,c))$ to $S_{\textit{\tiny rch}}$ where $R_1(a,b),R_2(b,c)\rightarrow R_3(a,c)\in P^*$,
    $R_1\circ R_2\sqsubseteq R_3\in\mathcal{R}$ and $R_2(b,c)\in V$.\hfill$\Box$
\end{enumerate}

Based on Algorithm~$\mathsf{PRC}$, we modify Algorithm~$\mathsf{A}_{opt}$
by replacing the step \ref{rch} of Algorithm~$\mathsf{OPT}$
to Algorithm~$\mathsf{PRC}$. Algorithm~$\mathsf{A}_{prc}$
is thus given as follows:\\

\noindent\texttt{Algorithm~$\mathsf{A}_{prc}$}. Given
a DHL$(\circ)$ ontology $\mathcal{O}=\langle\mathcal{T},\mathcal{R},\mathcal{A}\rangle$ and its
datalog program $P=\langle R, \textbf{I}\rangle$, this algorithm
returns a materialization graph $\mathcal{G}$ of $P$.
Initially $\mathcal{G}$ is empty. The following steps are then performed:
\begin{enumerate}[leftmargin=8ex,label=(\textit{Step \arabic*}),ref=Step~\arabic*]
\item Add all facts in $\textbf{I}$ to $\mathcal{G}$.\label{alg4:addFacts}
\item Compute $S_{\textit{\tiny rch}}$ by performing Algorithm~$\mathsf{PRC}$; use an NC
    algorithm to compute the transitive closure $S^*_{\textit{\tiny rch}}$ (see \ref{transClos} in Algorithm~$\mathsf{OPT}$);
    update $\mathcal{G}$ by performing \ref{updateG} in Algorithm~$\mathsf{OPT}$.\label{alg4:updateG}
\item If no node has been added to $\mathcal{G}$ (in \ref{alg4:updateG}), terminate,
    otherwise iterate \ref{alg4:updateG}. \label{alg4:halt}\hfill$\Box$
\end{enumerate}

\begin{theorem}\label{theorem:aprc}
For any DHL$(\circ)$ ontology $\mathcal{O}$, Algorithm~$\mathsf{A}_{prc}$ halts and outputs
a materialization graph of $\mathcal{O}$.
\end{theorem}

The above theorem is given to show the correctness of Algorithm~$\mathsf{A}_{prc}$.
We next use an example to show how Algorithm~$\mathsf{A}_{prc}$ handles
DHL$(\circ)$ materialization.

\begin{example}\label{exp:prc}
Consider performing Algorithm~$\mathsf{A}_{prc}$ on the ontology
$\mathcal{O}_{ex_1}$ in Example~\ref{exp:mg}. Note that the individual
set \textbf{IN}$=\{a_1,...,a_k,b\}$ where $k+1$ individuals are involved.
Initially, Algorithm~$\mathsf{A}_{prc}$ adds
all the facts ($A(b),R(a_1,b),S(a_2,a_1),...,S(a_{k},a_{k-1})$) to the
result $\mathcal{G}_{ex_1}$ (\ref{alg4:addFacts}). In the first iteration of
\ref{alg4:updateG}, Algorithm~$\mathsf{A}_{prc}$ computes $S_{\textit{\tiny rch}}$ first.
According to Algorithm~$\mathsf{PRC}$, for each atom of the form $S(a_i,a_{i-1})$ ($2\leq i\leq k$)
and $\forall o\in$\textbf{IN},
an $\texttt{\emph{rch}}$ relation of the form $\texttt{\emph{rch}}(R(a_{i-1},o),R(a_i,o))$ is added to
$S_{\textit{\tiny rch}}$. Since the atom $R(a_1,b)$ has been added to $\mathcal{G}_{ex_1}$,
Algorithm~$\mathsf{A}_{prc}$ checks that all atoms of the form $R(a_i,b)$ ($2\leq i\leq k$)
have SWD paths and are added to $\mathcal{G}_{ex_1}$ in the first iteration.
In the second iteration of \ref{alg4:updateG}, since all atoms of the form $R(a_i,b)$ ($1\leq i\leq k$)
have been in $\mathcal{G}_{ex_1}$, the $\texttt{\emph{rch}}$ relations of the form $\texttt{\emph{rch}}(A(b),A(a_i))$
are checked with respect to the axiom $\exists R.A\sqsubseteq A$; further all
atoms of the form $A(a_i)$ ($1\leq i\leq k$) are finally added to $\mathcal{G}_{ex_1}$
by Algorithm~$\mathsf{A}_{prc}$.
\end{example}

From the above example, one can find that Algorithm~$\mathsf{A}_{prc}$ terminates after two iterations
of \ref{alg4:updateG}. This is the same as Algorithm~$\mathsf{A}_{opt}$ (see Example~\ref{exp:opt}).
We use Theorem~\ref{theorem:aprcpt} to show that Algorithm~$\mathsf{A}_{prc}$ also has an NC version when handling ontologies
that follow the simple-concept and the simple-role restrictions.
The correctness of this theorem is based on that the method of narrowing down the scope of rule instantiations in Algorithm~$\mathsf{A}_{prc}$
does not influence the determination of the existence of SWD paths. The detailed analysis
can be found in the proof.

\begin{theorem}\label{theorem:aprcpt}
For any DHL$(\circ)$ ontology $\mathcal{O}$ that follows the simple-concept and the simple-role
restrictions,
there exists a poly-logarithmically bounded function $\psi$,
such that Algorithm~$\mathsf{A}_{prc}^{\psi}$ outputs
a materialization graph of $\mathcal{O}$.
\end{theorem}




%%% Local Variables:
%%% mode: latex
%%% TeX-master: "parallel-tractability-J"
%%% End:
